---
title: "Project 1"
author: "Tora Mullings"
date: "2/26/2022"
output:
  html_document: default
  pdf_document: default
---

# Load the necessary libraries.
readr, tidyverse and stringr.
```{r}
library(readr)
library(stringr)
library(tidyverse)
```

# Read file
Read the tournament text file.
```{r}
filename <- "tournamentinfo.txt"

rounds_line <- read_lines(filename, skip=2, n_max = 1)
raw_lines = read_lines(filename, skip=4, n_max = -1)
```

# Get the number of rounds in the tournament.
```{r}
str <- str_split(string=gsub("[^0-9]", "", rounds_line), pattern="")
round <- tail(unlist(str), n=1)
round <- as.numeric(round)
```


## `get_rounds_names` creates a vector of rounds names.
```{r}
get_round_names <- function(r) {
  names <- c()
  for (i in 1:r) {
    name <- paste(c("r", i), collapse="")
    names <- append(names, name)
  }
  return(names)
}

round_names <- get_round_names(round)
```


# `read_players`
`read_players` takes the raw lines from the text file. It initializes a data frame. Since each player in the text file has 2 lines, this function works 2 lines at a time. It stores the information from the text file in the data frame to be queried later.

```{r}
read_players <- function(lines) {
  df = data.frame()  # initialize dataframe
  i=1
  while (i < length(lines)) {
    player <- add_player(lines[i], lines[i+1])
    vec <- c(
      player_num=player["player_num"],
      state=player["state"],
      player_name=str_to_title(player["player_name"]),
      total_points=player["total_points"],
      player_pre_rating=player["player_pre_rating"])
    
    for (j in 1:round) {
      vec <- append(vec, player[round_names[j]])
    }
      
      
    df <- rbind(df, vec)
    
    i = i + 3   # skip --------------- and go to next player
  }
  columns= append(c("player_num","state","player_name", "total_points","player_pre_rating"),
                  round_names)
  colnames(df) = columns
  
  df[c(1,4:round+4+1)] <- as.numeric(unlist(df[c(1,4:round+4+1)]))
  
  return(df)
}
```



# `add_player`
`add_player` performs a series of regex operations on the pair of lines that belongs to each player. It removes whitespace and unnecessary characters. It returns vectors that hold clean data for a single player.
```{r}
add_player <- function(line1, line2) {
  # # parse line 1
  # 
  
  a <- str_split(line1, "[|]")
  a <- gsub('(\\s{2,})', '', a)
  a <- gsub('\\"', '', a)
  a <- unlist(str_split(a, ","))
  a <- gsub('(\\s{2,})', '', a)
  a[4:10] <- gsub("[A-Z]", "", a[4:10])
  a[1] <- gsub("c[(]", "", a[1])   # remove c(
  a <- a[1:length(a)-1]
  
  n <- c("player_num", "player_name", "total_points")
  n <- append(n, round_names)
  names(a) <- n
  
  
  # parse line 2
  b <- unlist(str_split(line2, "[|]"))
  b <- b[1:2]
  b <- gsub("\\s+", "", b)
  colon_index <- unlist(gregexpr(":", b))[2]
  arrow_index <- unlist(gregexpr("->", b))[2] 
  player_pre_rating <- substr(b, start=colon_index+1, stop=arrow_index-1)[2] # extract the pre-rating
  
  # Still might have a "P" in the score, remove it.
  player_pre_rating <- gsub("P.*", "", player_pre_rating)
  state <- b[1]
  vec <- c(state, player_pre_rating)
  names(vec) <- c("state", "player_pre_rating")
  
  return(c(a, vec))
}
```




# `get_opp_ids`
`get_opp_ids` takes a data frame of player information, and a player's id. It returns a vector of the player's opponents' ids.
```{r}
get_opp_ids <- function(df, id) {
  opps <- unlist(df[id,][c(6:(round+6-1))])
  return(opps)
}
```



# `get_opp_avg`
`get_opp_avg` takes a data frame of player information and a vector of player ids and returns the average of the corresponding players' pre-rating scores.
```{r}
get_opp_avg <- function(df, ids) {
  ratings <- c()
  for (player_id in ids) {
    ratings <- append(ratings, as.numeric(unlist(df[player_id,][5])))
    
  }
  avg <- mean(ratings, na.rm=TRUE)
  return(avg)
}
```



# Populate data frame
Place the cleaned data from the text file into a data frame named `result`.
```{r}
result <- read_players(raw_lines)
```


# Average Pre Chess Rating of Opponents
Calculate the pre chess rating averages of opponents for each player and place them into a vector. Append the vector as a new column in the data frame `result`.
```{r}
new_col <- c()
for (player in 1:nrow(result)) {  # player is a row
  opp_ids <- get_opp_ids(result, player)    # get ids of this player's opponents
  avg <- get_opp_avg(result, as.numeric(opp_ids))
  avg <- round(avg, digits=0)
  new_col <- append(new_col, avg)
}

result <- mutate(result, avg_opp_pre_rating=new_col)


```


# Drop columns
Drop the unnecessary columns and rearrange columns.
```{r}

result <- result[-c(1, 6:(round+6-1))]
View(result)
result <- result[c("player_name", "state", "total_points", "player_pre_rating", "avg_opp_pre_rating")]

```


# Data frame -> CSV file
Write the data frame to a csv file named `result.csv`.
```{r}
write.csv(result, "result.csv", row.names=FALSE)
```














